\Large \bf Evaluating eBPF as a Platform for Congestion Control Algorithm

### 2024-12-10 paper/sections/abstract.tex

\documentclass[../main.tex]{subfiles}
\begin{document}
Your abstract text goes here. Just a few facts. Whet our appetites.
Not more than 200 words, if possible, and preferably closer to 150.
\end{document}

>>>>

\documentclass[../main.tex]{subfiles}
\begin{document}
This paper seeks to explore the viability of developing Congestion Control Algorithms (CCAs) via eBPF. Developing new CCAs has traditionally been a time-consuming and error-prone process. Creating these algorithms directly within the Linux kernel demands a high level of kernel programming expertise while still leaving room for bugs that can crash the system. Although implementing CCAs as kernel modules has traditionally been an accepted alternative (https://pbg.cs.illinois.edu/papers/jay18pcc-kernel.pdf), it still runs the risk of introducing instability and crashes. Recent research has implemented a Congestion Control Plane (CCP) \cite{ccp} that allows CCA developers to implement CCAs in userspace, eliminating the need to run experimental software directly in the kernel. This approach involves a carefully designed and safety-checked datapath API; however, the Linux kernel datapath API still relies on a kernel module. Simultaneously, developing CCAs in eBPF \cite{ebpf_cca} has emerged as another promising approach, leveraging the eBPF verifier to ensure safety before the code ever runs in the kernel. In this paper, we compare these various approaches -- kernel-based, eBPF-based, and userspace (via CCP) -- along with a new hybrid userspace eBPF-based approach (ebpfccp) in terms of behavioral correctness and performance overhead.
\end{document}

### 2024-12-10 paper/sections/conclusion.tex

\documentclass[../main.tex]{subfiles}
\begin{document}
In order to fully evaluate performance of eBPF-based CCAs against alternative approaches, we looked at prior research aimed at reducing CCA-development overhead. Specifically, the Congestion Control Plane was developed to be an abstraction layer where developers can experiment with new CCA ideas without worrying about the intricacies of kernel networking stacks. Most notably, CCP decoupled data gathering and CCA computation. A datapath module gathers information regarding a TCP flow and sends it to user space via various channels, such as netlink sockets and unix pipes. A user space program then listens to the provided data and sends back any updates to be made to the TCP connection state. Since the computation lives in user space, developers are shielded from the potential of introducing kernel-crashing bugs as long as the underlying CCP datapath implementation is safe. The separation of data-gathering and computation also allows for further platform/OS independence. While traditional CCAs are tightly coupled with the kernel networking stack they are developed on top of, CCP allows for the same CCA to be run on different platforms with minimal changes. A CCP-based CCA can be directly ran on a completely different host, as long as there exists an existing datapath that implements the CCP standard API. \\ 
Due to CCP's datapath-reliant nature, the performance of any CCP-based congestion control will heavily depend on the implementation of the underlying datapath, as well as the channel used to communicate between user space and kernel space. In our investigations to test the performance of CCP-based CCAs, we decided to build an alternative datapath module that is implemented as an eBPF program rather than a kernel module. In addition to providing more safety guarantees, this implementation also allows us to test how the performance of CCP-based CCAs can vary depending on the underlying datapath implementation. \\

\end{document}

>>>>

\documentclass[../main.tex]{subfiles}
\begin{document}
Conclusion
\end{document}

### 2024-12-10 paper/sections/design.tex

\documentclass[../main.tex]{subfiles}
\begin{document}
Your abstract text goes here. Just a few facts. Whet our appetites.
Not more than 200 words, if possible, and preferably closer to 150.
\end{document}

>>>>

\documentclass[../main.tex]{subfiles}
\begin{document}

\begin{figure}[t]
\centering
    \includegraphics[width=\columnwidth]{img/ccp-traditional}
    \caption{CCP's conventional Linux kernel datapath, implemented as a kernel module, utilizes \texttt{libccp} to communicate with the CCP agent.
        Communication (1) is accomplished via Netlink sockets or a custom character device.
    }\label{fig:ccptraditional}
\end{figure}

Figure~\ref{fig:ccptraditional} illustrates the current CCP stack.
Notably, the CCP algorithm and agent both run in user space, while the datapath is implemented and run completely in kernel space.
Conventionally, CCP datapath program implementations such as Linux kernel's TCP stack, mTCP, and Google's QUIC all utilize \texttt{libccp} to interface with the CCP agent (Portus) \cite{ccp}.

\subsection{eBPF Implementation}

Due to the eBPF's intense verifier requirements and stipulations, running \texttt{libccp} within the eBPF program was deemed difficult.
Consequently, we create a user space layer between the CCP agent and the Linux kernel's TCP stack, called eBPFCCP.

eBPFCCP is written in Rust and runs \texttt{libccp} in user space to communicate with the CCP agent while also managing and communicating with the eBPF-based datapath program.
eBPFCCP's role is to collect primitive congestion signals from the eBPF program, send them over to the CCP agent, and execute instructions sent by the CCP agent to influence TCP behavior.
As shown in Figure~\ref{fig:ebpfccp}, eBPFCCP communicates with the CCP agent via Unix-domain sockets and interfaces with the eBPF datapath program using BPF hash maps and BPF ring buffers.

\begin{figure}[t]
\centering
    \includegraphics[width=\columnwidth]{img/ebpfccp}
    \caption{The eBPFCCP stack runs \texttt{libccp} in user space, communicating (1) with the CCP agent via Unix-domain sockets and interfacing with the eBPF datapath program via (2) BPF hash maps and (3) BPF ring buffers.
    }\label{fig:ebpfccp}
\end{figure}

On the eBPF datapath side, we use eBPF's \texttt{struct\_ops} API to attach congestion control-specific hooks to the TCP stack.
Each of these hooks operate on a per-flow basis.
When a new flow is created, the eBPF datapath program records the flow's information in a BPF ring buffer called \texttt{create\_conn\_events}.
This signals eBPFCCP to retrieve the new flow's information and pass it to the CCP agent.
Similarly, when a flow is terminated, the eBPF datapath program writes an event to the \texttt{free\_conn\_events} ring buffer, which eBPFCCP reads to perform flow cleanup actions.
As new ACKs arrive, the eBPF datapath program reads the TCP state and calculates primitive congestion signals, which are then written to the \texttt{signals} ring buffer, read by eBPFCCP, and sent to the CCP agent.
Then, the message the CCP agent sends back to the eBPFCCP pertaining to window size and pacing rate is written to the \texttt{connections} hash map, which the eBPF datapath program reads and updates the TCP state accordingly.
The BPF maps used by the eBPF datapath and eBPFCCP are summarized in Table~\ref{tab:maptable}.

\begin{table}[htbp]
    \centering
    \footnotesize
    \begin{tabular}{@{}p{0.35\columnwidth} p{0.5\columnwidth}@{}}
        \toprule
        \textbf{Map} & \textbf{Description} \\[3pt]
        \midrule
        \texttt{connections} (BPF Hash) &
        eBPFCCP writes per-flow congestion window size and pacing rate. 
        The eBPF datapath reads these values and updates the TCP state accordingly. \\[6pt]

        \texttt{signals} (BPF Ringbuf) &
        The eBPF datapath writes primitive signal measurements into this buffer. 
        The eBPFCCP reads these measurements and forwards them to the CCP agent. \\[6pt]

        \texttt{create\_conn\_events} (BPF Ringbuf) &
        Whenever a new flow is created, the eBPF datapath records flow information here. 
        The eBPFCCP then retrieves and passes these details to the CCP agent. \\[6pt]

        \texttt{free\_conn\_events} (BPF Ringbuf) &
        Upon flow termination, the eBPF datapath writes an event. 
        The eBPFCCP reads and relays this information to the CCP agent for flow cleanup actions. \\[3pt]
        \bottomrule
    \end{tabular}
    \caption{BPF Maps used by the eBPF datapath and eBPFCCP for managing flow states, signals, and events.}
    \label{tab:maptable}
\end{table}

Overall, the eBPFCCP aims to act as a bridge between the CCP agent and the eBPF datapath program.

\end{document}
